Ornaments as indicators of social change before and after European arrival at Kiwulan, Northeastern Taiwan

### 2019-11-22 analysis/paper/paper.Rmd
updates based on reviews

This study focuses on `r sam_post_hole` ornaments from 40 sampling squares that were excavated _in situ_ from the household area of Kiwulan, indicated by numerous post-holes (Figure \@ref(fig:KWL-sam-area)). Thirty burials contained 3,173 ornaments, and 27 ornaments were recovered from middens. In the burials ornaments were in the form of bead strings that consisted of hundreds to thousands of pieces per burial. The ornaments from the burial contexts are not statistically analysed because most burials were assigned to 17-18 th century that offers limited quantitative comparison with other time periods. But the use of ornaments as body adornments in burials is discussed to understand the functions of ornaments. The household area represents domestic spaces suitable for testing for changes in ornament distribution relating to social organisation. Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), agate beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)). The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).

>>>>

This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aliened postholes with in-situ posts (Figure \@ref(fig:KWL-sam-area)). These postholes were found aligned in a north-south direction in some intervals with construction marks, which were interpreted as the remains of silt house structures. At the north margin of the dwelling place were burials that are mostly oriented in an east-west direction across the sampling area. There are 30 burials in the sampling area account for one thrid of the total number of burials at Kiwulan. Ornaments are commonly used as grave goods in burials, and the total number including different ornament classes are 3,173. The use of ornaments as body adornments in burials will be discussed to understand the functions of ornaments, but not statistically analysed since the number of ornaments is highly affected by the presence of bead strings, which sometimes contains thousands of beads in an individual burial [@Chen2007]. In addition to ornaments from the dwelling place and burial, there are 27 ornaments found in middens, which to some extent supports that ornaments were treated as high value items at Kiwulan because of small number in the midden contexts compared to high frequency in burial contexts. It is important to analyse the detailed distribution of ornaments from the dwelling place as it can provide key information concerning the spatial pattens across the space as well as changes in ornament relating to social organisation across different time periods.

The ornaments found in the  Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament are golden beads (n = `r golden_bead_n`), followed by glass beads (n =  `r glass_bead_n`), agate beads (n = `r agate_bead_n`), metal rings (n = `r metal_ring_n`), and metal bells (n = `r metal_bell_n`, see also Figure \@ref(fig:ornaments-image)). The variety of subtypes and their individual frequencies are summarized in Table \@ref(tab:tidy-data-for-table).
